---
title: "Project Milestone 4"
format:
  html:
    theme: flatly
    mainfont: "system-ui, -apple-system, 'Segoe UI', Roboto, 'Helvetica Neue', Arial, sans-serif"
    page-layout: full
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(plotly)
library(htmlwidgets)
```

:::: {.columns}

::: {.column width="68%"}

```{r}
#| label: animal-breeds-plot
#| message: false

# Load sample open animal shelter data
shelter_data <- read_csv(
  "https://data.austintexas.gov/resource/wter-evkm.csv?$limit=2000",
  show_col_types = FALSE
)

# Find top 15 most common breeds
top_breeds <- shelter_data |>
  count(animal_type, breed, sort = TRUE) |>
  slice_head(n = 15)

# Calculate overall counts and percentages for each animal type
stats_list <- shelter_data |>
  count(animal_type) |>
  mutate(pct = round(n / sum(n) * 100, 1)) |>
  split(~animal_type) |>
  lapply(\(df) list(count = df$n, pct = df$pct))

stats_json <- jsonlite::toJSON(stats_list, auto_unbox = TRUE)

# Create interactive bar plot showing breeds colored by animal type
p <- top_breeds |>
  ggplot(aes(x = reorder(breed, n), y = n, fill = animal_type)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Top Animal Breeds Sent to Shelters",
    x = "Breed",
    y = "Count",
    fill = "Animal Type"
  ) +
  theme_minimal()

# Connect graph clicks to the summary element below
js_code <- sprintf("
function(el, x) {
  var stats = %s;
  function updateSummary(animalType, breed, count) {
    var summaryEl = document.getElementById('animal-summary');
    if (!summaryEl || !stats[animalType]) return;
    var s = stats[animalType];
    var breedInfo = breed ? ('You selected <strong>' + breed + '</strong> (' + count.toLocaleString() + ' animals).<br>') : '';
    summaryEl.innerHTML = '<h4>' + animalType + ' Summary</h4>' +
      '<p>' + breedInfo +
      'Across all recorded shelter intakes, <strong>' + animalType + 's</strong> make up <strong>' +
      s.count.toLocaleString() + '</strong> animals, which is <strong>' + s.pct + '%%%%</strong> of all animals sent to the shelter.</p>';
  }

  el.on('plotly_click', function(d) {
    if (!d || !d.points || d.points.length === 0) return;
    var pt = d.points[0];
    updateSummary(pt.data.name, pt.y, pt.x);
  });

  el.on('plotly_legendclick', function(d) {
    var animalType = d.data[d.curveNumber].name;
    updateSummary(animalType, null, null);
  });
}
", stats_json)

ggplotly(p) |>
  onRender(js_code)
```

::: {#animal-summary style="margin-top: 15px; margin-bottom: 25px; padding: 15px; background-color: #f8f9fa; border-left: 5px solid #0d6efd; border-radius: 4px;"}
### Animal Selection Summary
*Click any animal breed bar or legend item in the graph above to view its percentage breakdown in comparison to all animals.*
:::

### Overall Intake Breakdown by Animal Type

```{r}
#| label: summary-table

shelter_data |>
  count(animal_type, name = "Total Intakes") |>
  mutate(`Percentage of All Animals` = sprintf("%.1f%%", `Total Intakes` / sum(`Total Intakes`) * 100)) |>
  arrange(desc(`Total Intakes`)) |>
  knitr::kable()
```

:::

::: {.column width="4%"}
:::

::: {.column width="28%"}

::: {.card .p-3 style="background-color: #f8f9fa; border: 1px solid #dee2e6; border-radius: 8px;"}
### 🏛️ Top 10 States for Shelter Intakes

Based on national reporting from Shelter Animals Count and Best Friends Animal Society:

| Rank | State | Est. Intakes |
|:----:|:------|-------------:|
| 1 | Texas | ~680,000 |
| 2 | California | ~520,000 |
| 3 | Florida | ~370,000 |
| 4 | North Carolina | ~220,000 |
| 5 | Georgia | ~190,000 |
| 6 | Ohio | ~150,000 |
| 7 | Louisiana | ~130,000 |
| 8 | New York | ~125,000 |
| 9 | Illinois | ~115,000 |
| 10 | Arizona | ~110,000 |

*These 10 states account for more than 50% of all shelter intakes nationwide.*
:::

:::

::::
